---
title: "Investigation of Floods (Manuel Segment)"
author: "Team - Awesome"
css: styles.css
output: html_document
---

```{r libraries, echo=F, message=F, warning=F}
library(lubridate)
library(animation)
library(reshape2)
library(RColorBrewer)
library(gridExtra)
library(ggmap)
library(dplyr)
```

```{r dataClean, echo=F, message=F, warning=F}
################
## DATA CLEAN ##
################

## Read floods.
floods <- read.csv("GlobalFloodsRecord.csv", stringsAsFactors = F)

## Clean and generate new columns.
floods$lon <- as.numeric(floods$Centroid.X)
floods$lat <- as.numeric(floods$Centroid.Y)
floods$dead <- as.numeric(floods$Dead)
floods$displaced <- as.numeric(floods$Displaced)
floods$severity <- as.numeric(floods$Severity..)
floods$damage <- as.numeric(floods$Damage..USD.)
floods$country <- as.factor(floods$Main.cause)
floods$other <- as.factor(floods$Country)
floods$cause <- as.factor(floods$Other)
floods$magnitude <- as.numeric(floods$Magnitude..M...)
floods$began <- as.Date(floods$Began,"%d-%b-%y")
floods$ended <- as.Date(floods$Ended,"%d-%b-%y")
floods$start <- year(floods$began)
floods$end <- year(floods$ended)


## Remove useless columns and add grouping dates.
floods2 <- tbl_df(floods[,-(1:30)])
floods2 <- floods2[!is.na(floods2$began),]
breaks = round(seq(1985, 2016, 7.75))
floods2$breaks <- cut(floods2$start, breaks = breaks, labels = breaks[-1], right = F)
levels(floods2$breaks) <- c("1985-1993","1994-2000","2001-2008","2009-2016")
```

As a next step in the analysis, we look at the distribution of floods across the different continents throughout time. The goal is to identify possible patterns, and address questions such as:

- Have floods concentrated on particular continents during specific periods?
- For any given continent, have there been more disastrous times in terms of floods and displacement?
- Within continents, which regions are more succeptible to disasters caused by floods?

To do this, 

```{r byContinent, echo=F, message=F, warning=F, fig.width= 13, fig.height = 8}
#########################
## FLOODS BY CONTINENT ##
#########################

flood.cols <- brewer.pal(n = 5, name = "OrRd")

## NORTH AMERICA
america <- get_map("America", zoom =4 , maptype = "roadmap")

america.map <- ggmap(america,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "North America") +
  theme(legend.position = "none") +
  coord_map()

america.facet <- america.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## SOUTH AMERICA
south.america <- get_map("South America", zoom =4 , maptype = "roadmap")

south.america.map <- ggmap(south.america,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "South America") +
  theme(legend.position = "none") +
  coord_map()

south.america.facet <- south.america.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## PLOT THESE 2
grid.arrange(america.facet, south.america.facet, ncol = 2)

## EUROPE
europe <- get_map("Europe", zoom =4 , maptype = "roadmap")

europe.map <- ggmap(europe,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "Europe") +
  theme(legend.position = "none") +
  coord_map()

europe.facet <- europe.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## ASIA
asia <- get_map("Asia", zoom =4 , maptype = "roadmap")

asia.map <- ggmap(asia,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "Asia") +
  theme(legend.position = "none") +
  coord_map()

asia.facet <- asia.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## PLOT THESE 2
grid.arrange(europe.facet, asia.facet, ncol = 2)

## AFRICA
africa <- get_map("Africa", zoom =4 , maptype = "roadmap")

africa.map <- ggmap(africa,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "Africa") +
  theme(legend.position = "none") +
  coord_map()

africa.facet <- africa.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## AUSTRALIA
australia <- get_map("Australia", zoom =4 , maptype = "roadmap")

australia.map <- ggmap(australia,  extent="panel") %+%
  floods2 +
  #aes(x = lon, y = lat, z = log(displaced)) +
  aes(x = lon, y = lat) +
  #stat_summary_2d(fun = "sum", bins = 60, alpha = 0.75) + 
  stat_density2d(aes(fill = ..level.., alpha = ..level..), geom = "polygon") +
  scale_fill_gradientn(name = "Floods", colours = flood.cols, space = "Lab") +
  scale_alpha(range = c(0.1, 0.5)) +
  labs(x = "Longitude", y = "Latitude", title = "Australia") +
  theme(legend.position = "none") +
  coord_map()

australia.facet <- australia.map +
  geom_point(aes(x = lon, y = lat, size = displaced), fill="red", shape=21, alpha=0.8) +
  facet_wrap(~ breaks)

## PLOT THESE 2
grid.arrange(africa.facet, australia.facet, ncol = 2)

```