---
title: "map part"
author: "Yusen Wang"
date: "February 27, 2016"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ncdf4)
library(xlsx)
library(reshape)
library(dplyr)
library(ggplot2)
fname <- 'new_phi_data.nc'
ncin <- nc_open(fname)

time <- ncvar_get(ncin, 'T')
tunits <- ncatt_get(ncin, 'T', 'units')

lon <- ncvar_get(ncin,'X')
lat <- ncvar_get(ncin, 'Y')

pre <- ncvar_get(ncin, 'P')

phi.array <- ncvar_get(ncin, 'phi')

year = 1948
year_total= 0
total_35  = 0 
total_70 = 0
days = 0
year_ave = NULL
ave_35_ls = NULL
ave_70_ls = NULL
for (i in 0:(length(time)-1)){
  this_date = as.Date(i,origin = '1948-01-01')
  this_year = as.numeric(format(this_date,'%Y'))
  day_ave = mean(phi.array[,,i+1])
  ave_35 = mean(phi.array[,15:25,i+1])
  ave_70 = mean(phi.array[,1:14, i+1])
  if (this_year==year){
      year_total = year_total + day_ave
      total_35 = total_35 + ave_35
      total_70 = total_70 +ave_70
      days = days + 1
  }
  else{
    year_mean = year_total/days
    mean_35 = total_35/days
    mean_70 = total_70/days
    year_ave = append(year_ave, year_mean)
    ave_35_ls = append(ave_35_ls, mean_35)
    ave_70_ls = append(ave_70_ls,mean_70)
    year = year + 1
    year_total = day_ave
    total_35 = ave_35
    total_70 = ave_70
    days = 1
  }
  
}
year_mean = year_total/days
year_ave = c(year_ave, year_mean)
mean_35 = total_35/days
mean_70 = total_70/days
ave_35_ls = append(ave_35_ls, mean_35)
ave_70_ls = append(ave_70_ls, mean_70)

phi_year = data.frame((1948 + 0:(length(year_ave)-2)),year_ave[-length(year_ave)]) 
colnames(phi_year) = c('year', 'ave_phi')

phi_year_35 = data.frame((1948 + 0:(length(ave_35_ls)-2)),ave_35_ls[-length(ave_35_ls)]) 
colnames(phi_year_35) = c('year', 'ave_phi_35')

phi_year_70 = data.frame((1948 + 0:(length(year_ave)-2)),ave_70_ls[-length(ave_70_ls)]) 
colnames(phi_year_70) = c('year', 'ave_phi_70')

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
flood <- read.csv("GlobalFloodsRecord.csv", header = T)
nameuse <- c("Country", "Detailed.Locations..click.on.active.links.to.access.inundation.extents.", "Began", "Ended", "Duration.in.Days", "Dead", "Displaced", "Main.cause", "Severity..", "Affected.sq.km", "Magnitude..M...", "Centroid.X", "Centroid.Y", "M.6", "M.4")
floodinfo <- flood[ , nameuse]
colnames(floodinfo) <- c("Country", "Detailed.Locations", "Began", "Ended", "Duration", "Dead", "Displaced", "Cause", "Severity", "Affected.sq.km", "Magnitude", "Longitude", "Latitude", "Extreme", "Large")

floodinfo$Longitude <- as.numeric(as.character(floodinfo$Longitude))
floodinfo$Latitude <- as.numeric(as.character(floodinfo$Latitude))
floodinfo <- na.omit(floodinfo)

floodinfo$Began <- as.Date(as.character(floodinfo$Began), "%d-%B-%y")
floodinfo$Ended <- as.Date(as.character(floodinfo$Ended), "%d-%B-%y")

floodinfo_N35_N70 <- filter(floodinfo, Latitude <= 70 & Latitude >= 35)
floodinfo_N10_N35 <- filter(floodinfo, Latitude < 35 & Latitude >= 10)

############################################# World
annual_num_extreme_flood_yr <- floodinfo %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Extreme) %>% 
        summarise(n()) %>% filter(Extreme == 1)
        
annual_num_large_flood_yr <- floodinfo %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Large) %>% 
        summarise(n()) %>% filter(Large == 1)
annual_num_large_flood_yr <- annual_num_large_flood_yr[-1, ]

bind = data.frame(annual_num_large_flood_yr$Began_yr, annual_num_large_flood_yr$`n()`)
colnames(bind) = c('year','large')
ggplot(phi_year, aes(x = year, y = ave_phi))+ geom_point(size = 3, alpha = 0.5, color = '#66b3ff')+geom_smooth(se = F,size = 0.75, color = '#3399ff', linetype = 'dashed') +theme_bw()+ ggtitle('Average Phi by Years') + ylab('Average Phi') 

ggplot(data = bind, aes(x = year, y = large))+geom_line(size = 3, alpha = 0.5, color = '#66b3ff')+theme_bw()+ ggtitle('M4') + ylab('# of floods') 

bind2 = data.frame(annual_num_extreme_flood_yr$Began_yr, annual_num_extreme_flood_yr$`n()`)
colnames(bind2) = c('year','extreme')
ggplot(data = bind2, aes(x = year, y = extreme))+geom_line(size = 3, alpha = 0.5, color = '#ff6600')+theme_bw()+ ggtitle('M6') + ylab('# of floods')
        
################################################### N35-N70
annual_num_extreme_flood_70 <- floodinfo_N35_N70 %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo_N35_N70$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Extreme) %>% 
        summarise(n()) %>% filter(Extreme == 1)
        
annual_num_large_flood_70 <- floodinfo_N35_N70 %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo_N35_N70$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Large) %>% 
        summarise(n()) %>% filter(Large == 1)
annual_num_large_flood_yr <- annual_num_large_flood_70[-1, ]

bind = data.frame(annual_num_large_flood_70$Began_yr, annual_num_large_flood_70$`n()`)
colnames(bind) = c('year','large')
ggplot(phi_year_70, aes(x = year, y = ave_phi_70))+ geom_point(size = 3, alpha = 0.5, color = '#66b3ff')+geom_smooth(se = F,size = 0.75, color = '#3399ff', linetype = 'dashed') +theme_bw()+ ggtitle('Average Phi by Years_N35-N70') + ylab('Average Phi') 

ggplot(data = bind, aes(x = year, y = large))+geom_line(size = 3, alpha = 0.5, color = '#66b3ff')+theme_bw()+ ggtitle('M4') + ylab('# of floods') 

bind2 = data.frame(annual_num_extreme_flood_70$Began_yr, annual_num_extreme_flood_70$`n()`)
colnames(bind2) = c('year','extreme')
ggplot(data = bind2, aes(x = year, y = extreme))+geom_line(size = 3, alpha = 0.5, color = '#ff6600')+theme_bw()+ ggtitle('M6') + ylab('# of floods')


################################################## N10-N35
annual_num_extreme_flood_35 <- floodinfo_N10_N35 %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo_N10_N35$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Extreme) %>% 
        summarise(n()) %>% filter(Extreme == 1)
        
annual_num_large_flood_35 <- floodinfo_N10_N35 %>%
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo_N10_N35$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Large) %>% 
        summarise(n()) %>% filter(Large == 1)
annual_num_large_flood_35 <- annual_num_large_flood_35[-1, ]

bind = data.frame(annual_num_large_flood_35$Began_yr, annual_num_large_flood_35$`n()`)
colnames(bind) = c('year','large')
ggplot(phi_year_35, aes(x = year, y = ave_phi_35))+ geom_point(size = 3, alpha = 0.5, color = '#66b3ff')+geom_smooth(se = F,size = 0.75, color = '#3399ff', linetype = 'dashed') +theme_bw()+ ggtitle('Average Phi by Years_N10-N35') + ylab('Average Phi') 

ggplot(data = bind, aes(x = year, y = large))+geom_line(size = 3, alpha = 0.5, color = '#66b3ff')+theme_bw()+ ggtitle('M4') + ylab('# of floods') 

bind2 = data.frame(annual_num_extreme_flood_35$Began_yr, annual_num_extreme_flood_35$`n()`)
colnames(bind2) = c('year','extreme')
ggplot(data = bind2, aes(x = year, y = extreme))+geom_line(size = 3, alpha = 0.5, color = '#ff6600')+theme_bw()+ ggtitle('M6') + ylab('# of floods')


###################################################       
        
        #plot(phi_year_plot, annual_num_extreme_flood$`n()`, xlab = "Pressure by year: 1985-2015", ylab = "Total Number of Annual Extreme Floods", main = "The Relation Between Pressure and Number of Annual Extreme Floods")
       # abline(lm(annual_num_extreme_flood$`n()` ~ phi_year_plot))
       # plot(phi_year_plot, annual_num_large_flood$`n()`, xlab = "Pressure by year: 1985-2015", ylab = "Total Number of Annual Large Floods", main = "The Relation Between Pressure and Number of Annual Large Floods")
       # abline(lm(annual_num_large_flood$`n()` ~ phi_year_plot))
###########################################################################
```

We can see that total annual floods have an apparent periodicity. Large floods and extreme floods share a similar trend. What should be concerned is that we are in a dangerous time period now, because in 2015, large floods seem to touch the bottom of its periodicity. Maybe there would be more floods in 2016. 

```{r,results='asis',tidy=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')

floodinfo$Latlong <- paste(floodinfo$Latitude, floodinfo$Longitude, sep = ":")
FloodMap <- gvisMap(floodinfo, "Latlong", "Began",
                     options=list(showTip=TRUE, 
                                  showLine=TRUE, 
                                  enableScrollWheel=TRUE,
                                  mapType='terrain', 
                                  useMapTypeControl=TRUE))
plot(FloodMap)
```


Compare with the dynamic map which is created in shiny app, https://yusenwang.shinyapps.io/Dynamic_Flood_Map/
We can see that .......

We can know from these pictures that pressure has a significant linear effect on number of floods. For extreme floods, it has less variance, while more variance on large floods. 



```{r}

floodinfo_N35_N70$Severity <- as.numeric(as.character(floodinfo_N35_N70$Severity))
for (j in 1:length(floodinfo_N35_N70)) {
        if (floodinfo_N35_N70$Longitude[j] < 0) {
                floodinfo_N35_N70$Longitude[j] <- floodinfo_N35_N70$Longitude[j] + 360
        }
}



extract_info <- function(floodinfor_N35_N70, phi.array) {
        phi_sample <- rep(NA, nrow(floodinfo_N35_N70))
        for (i in 1:nrow(floodinfo_N35_N70)) {
                time_min <- as.numeric(floodinfo_N35_N70$Began[i]
                                       -as.Date("1948-01-01"))-15
                time_max <- as.numeric(floodinfo_N35_N70$Ended[i]
                                       -as.Date("1948-01-01"))+15
                
                lat_min <- floodinfo_N35_N70$Latitude[i] - 2.5
                lat_min_index <- ceiling((lat_min-35) / 2.5)
                lat_max <- floodinfo_N35_N70$Latitude[i] + 2.5
                lat_max_index <- floor((lat_max-35) / 2.5)
                
                lon_min <- floodinfo_N35_N70$Longitude[i] - 2.5
                lon_min_index <- ceiling(lon_min / 2.5)
                lon_max <- floodinfo_N35_N70$Longitude[i] + 2.5
                lon_max_index <- floor(lon_max / 2.5)
                
                phi_sample[i] <- mean(phi.array[lon_min_index:lon_max_index, lat_min_index:lat_max_index, time_min:time_max])
        }
        return(phi_sample)
}

pressure_level <- extract_info(floodinfo_N35_N70, phi.array)

#lm(flood_N35_N70$Severity~pressure_level)
#plot(flood_N35_N70$Severity~pressure_level)
#pc1 <- princomp(phi.array, cor = TRUE, scores = TRUE)
#summary(pc1)

```
