---
title: "map part"
author: "Yusen Wang"
date: "February 27, 2016"
output: html_document
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
flood <- read.csv("GlobalFloodsRecord.csv", header = T)

nameuse <- c("Country", "Detailed.Locations..click.on.active.links.to.access.inundation.extents.", "Began", "Ended", "Duration.in.Days", "Dead", "Displaced", "Main.cause", "Severity..", "Affected.sq.km", "Magnitude..M...", "Centroid.X", "Centroid.Y", "M.6", "M.4")
floodinfo <- flood[ , nameuse]
colnames(floodinfo) <- c("Country", "Detailed.Locations", "Began", "Ended", "Duration", "Dead", "Displaced", "Cause", "Severity", "Affected.sq.km", "Magnitude", "Longitude", "Latitude", "Extreme", "Large")

floodinfo$Longitude <- as.numeric(as.character(floodinfo$Longitude))
floodinfo$Latitude <- as.numeric(as.character(floodinfo$Latitude))
floodinfo <- na.omit(floodinfo)

floodinfo$Began <- as.Date(as.character(floodinfo$Began), "%d-%B-%y")
floodinfo$Ended <- as.Date(as.character(floodinfo$Ended), "%d-%B-%y")

annual_num_extreme_flood <- floodinfo %>% 
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Extreme) %>% 
        summarise(n()) %>% filter(Extreme == 1)
 
annual_num_large_flood <- floodinfo %>% 
        mutate(Began_yr=as.numeric(format(as.POSIXct(floodinfo$Began,format="%Y-%m-%d %I:%M:%S %p"),"%Y"))) %>% 
        group_by(Began_yr, Large) %>% 
        summarise(n()) %>% filter(Large == 1)
annual_num_large_flood <- annual_num_large_flood[-1, ]

op <- par()
par(mfrow = c(2, 1))
plot(annual_num_large_flood$Began_yr, annual_num_large_flood$`n()`, "l", xlab = "Year", ylab = "Number", las = 3)
title("Total Annual Large Floods (M>4) ")
plot(annual_num_extreme_flood$Began_yr, annual_num_extreme_flood$`n()`, "l", xlab = "Year", ylab = "Number", las = 3)
title("Total Annual Extreme Floods (M>6)")
par(op)
```

We can see that total annual floods have an apparent periodicity. Large floods and extreme floods share a similar trend. What should be concerned is that we are in a dangerous time period now, because in 2015, large floods seem to touch the bottom of its periodicity. Maybe there would be more floods in 2016. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(googleVis)

floodinfo$Latlong <- paste(floodinfo$Latitude, floodinfo$Longitude, sep = ":")
FloodMap <- gvisMap(floodinfo, "Latlong", "Began",
                     options=list(showTip=TRUE, 
                                  showLine=TRUE, 
                                  enableScrollWheel=TRUE,
                                  mapType='terrain', 
                                  useMapTypeControl=TRUE))
plot(FloodMap)
```
See the following link:
http://127.0.0.1:17543/custom/googleVis/MapID952b50f1f564.html

Compare with the dynamic map which is created in shiny app, https://yusenwang.shinyapps.io/Dynamic_Flood_Map/
We can see that .......

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Combine Cheng's part in this chunk.
library(ncdf4)
library(xlsx)
library(reshape)
fname <- 'NOAA_Daily_phi_500mb.nc'
ncin <- nc_open(fname)
#print(ncin)
#data = read.xlsx('GlobalFloodsRecord.xls',1)

time <- ncvar_get(ncin, 'T')
tunits <- ncatt_get(ncin, 'T', 'units')

lon <- ncvar_get(ncin,'X')
lat <- ncvar_get(ncin, 'Y')

pre <- ncvar_get(ncin, 'P')

phi.array <- ncvar_get(ncin, 'phi')

# worldwide
ave <- NULL
for (i in 1:24836){
  ave[i] <- mean(phi.array[,,i])
}
ave_yr <- colMeans(matrix(ave, nrow=365))

pre_yr <- ave_yr[38:(length(ave_yr)-1)]

plot(pre_yr, annual_num_extreme_flood$`n()`, xlab = "Pressure by year: 1985-2015", ylab = "Total Number of Annual Extreme Floods", main = "The Relation Between Pressure and Number of Annual Extreme Floods")
abline(lm(annual_num_extreme_flood$`n()` ~ pre_yr))

plot(pre_yr, annual_num_large_flood$`n()`, xlab = "Pressure by year: 1985-2015", ylab = "Total Number of Annual Large Floods", main = "The Relation Between Pressure and Number of Annual Large Floods")
abline(lm(annual_num_large_flood$`n()` ~ pre_yr))
```

We can know from these pictures that pressure has a significant linear effect on number of floods. For extreme floods, it has less variance, while more variance on large floods. 